---
title: ''
author: "Matthew R.V. Ross"
output:
  pdf_document:
    latex_engine: pdflatex
    template: rap-latex-cv-sketch.tex
  word_document: default
  html_document:
    df_print: paged
fontfamily: palatino
fontsize: 10pt
geometry: margin=1in
fontawesome: yes
urlcolor: blue
header-includes:
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{multirow} 
  \usepackage{wrapfig}
  \usepackage{float}
  \usepackage{colortbl}
  \usepackage{pdflscape}
  \usepackage{tabu}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage[normalem]{ulem}
  \usepackage{makecell}
  \usepackage{xcolor}
---

# Professional Preparation

```{r, include=F,echo=F}
library(knitr)
library(tidyverse)
library(kableExtra)
library(knitr)
```


```{r, echo=F,warning=F}
inst <- tibble(institution = c('University of Colorado',
                               'Duke University',
                               'University of North Carolina'),
               place=c('Boulder CO','Durham NC','Chapel Hill NC'),
               degree=c('B.A. Ecology and Evolultionary Biology, Minor in French','PhD Ecology',
                        'Post-doc in remote sensing'),
               year=c('2010','2017','2017-2018'))

options(knitr.table.format = 'latex')
names(inst) <- NULL
kable(inst,booktabs=T,longtable=T) %>%
  kable_styling(full_width=T,position='right') %>%
  column_spec(3,width='6cm')%>%
  column_spec(1,width='5cm') %>%
  column_spec(4,width='1.5cm')

  

  
```

*University of Colorado* \quad Boulder CO \quad   B.A. Ecology and Evol. Biology Minor in French \hfill 2010

*Duke University*    \quad \quad  \quad   Durham NC \quad \quad Ph.D. Ecology \hfill 2017

*University of North Carolina at Chapel Hill*; NC,  Post-doc in aquatic remote sensing \hfill 2017-2018

# Employment

*Colorado State University Department of Ecosystem Science and Sustainability*

 > Assistant professor of water quality \hfill 2018--present


# Publications

## 5 most relevant to this proposal 

AA Peri\c{c}ak CJ Thomas, DA Kroodsma, MF Wasson, MRV Ross, NE Clinton, DJ Campagna, Y Franklin, ES Bernhardt, JF Amos. *Mapping the yearly extent of surface coal mining in Central Appalachia using Landsat and Google Earth Engine*. PLOS ONE. 2018. 10.1371/journal.pone.0197758

MRV Ross, BL McGlynn, ES Bernhardt. *Deep impact: effects of mountaintop mining on surface topography, bedrock structure, and downstream waters*. 2016. Environmental science & technology. 10.1021/acs.est.5b04532

MRV Ross, F Nippgen, BA Hassett, BL McGlynn, ES Bernhardt. *Pyrite oxidation drives exceptionally high weathering rates and geologic $CO_2$ release in mountaintop-mined landscapes*. Global Biogeochemical Cycles. 2018. 10.1029/2017GB005798

F Nippgen, MRV Ross, ES Bernhardt, BL McGlynn. *Creating a more perennial problem? Mountaintop removal coal mining enhances and sustains saline baseflows of Appalachian watersheds*. Environmental science & technology. 2017. 10.1021/acs.est.7b02288


MRV Ross, ES Bernhardt, MW Doyle, JB Heffernan. *Designer ecosystems: incorporating design approaches into applied ecology*. 2015. Annual Reviews of Environment and Resources. 10.1146/annurev-environ-121012-100957

## Additional publications

Barger, N. N., Gardner, T. A., Sankaran, M., Belnap, J., Broadhurst, L., Brochier, V., 
Isbell, F., Meyfroidt, P., Moreira, F., Nieminen, T. M., Okuro, T., Rodrgiues, R. R., Saxena, V., and Ross, M. *Chapter 3: Direct and indirect drivers of land degradation and restoration*. In IPBES (2018): The IPBES assessment report on land degradation and restoration. Montanarella, L., Scholes, R., and Brainich, A. (eds.). Secretariat of the Intergovernmental Science-Policy Platform on Biodiversity and Ecosystem Services, Bonn, Germany, pp. 137-218


MC Arnold, LA Friedrich, TT Lindberg, MRV Ross, NM Halden, ES Bernhardt, VP Palace, RT Di Giullio. *Microchemical analysis of selenium in otoliths of two West Virginia fishes captured near mountaintop removal coal mining operations*. 2015. Environmental Toxicology and Chemistry. 10.1002/etc.2885


MR Ross, SC Castle, NN Barger. *Effects of fuels reductions on plant communities and soils in a Pi\~non Juniper woodland*. 2012. Journal of arid environments. 10.1016/j.jaridenv.2011.11.019

# Synergistic activities 

**Maintenance and construction of AquaSat** - AquaSat is a dataset that uses Google
Earth Engine and R to join *in-situ* measurements of Water Quality with Landsat
imagery. These matchups can be used to build continental scale predictions of
water quality. I am the lead maintainer and developer for this 
semi-annually updated dataset which is pending publication. 

**Lead for watershed major curriculum changes at CSU** -  I am leading changes 
in our curriculum at Colorado State to make sure our Watershed major students
have modern analytical skills. This involves a 
committment to teaching both faculty and students a cohesive approach to analyzing
watershed data and building tutorials to that end. 

**NREL Education Committee Chair** - As part of my position at CSU, I am the chair
of the Natural Resource Ecology Lab's educational activities. This includes activities
building, supporting, and maintaining summer education institutes to 
develop modern data analytics skills and share 
these analyses with a broad audience using the interactive web building tools
from Shiny R. 

**Collaborator with the USGS Integrated Information Dissemination Division** - 
For ongoing work and pending proposal's I am a close collaborator with the USGS
IIDD, part of the Water Mission Area at USGS. IIDD works to analyze, visualize,
and disseminate relavent water data to the world, and I have been involved with 
their efforts to streamline analysis pipelines as well as distribution of these
data. 

**Coordinator and lead of Open Eco Data Lab** - I am working with graduate
students in the Graduate Degree Program in Ecology at CSU to build a permanent 
and stable portal for students to learn how to use R. This portal will house both
tutorials built by other researchers and ones that I, my lab, or other students
create. 

